Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-07-10 doc/reviews-response-draft.txt

Happily, Reviewers #2 and #3's concerns about evaluation may be answered in the supplementary material. (You were not required to review the supplementary material, but we can point to it in our author response.)  We do not think that there is an evaluation gap in the paper, except for Reviewer #1's excellent suggestion that we could also compare with other semantic dependency representations (which may however be non-planar, disconnected, or not manually checked).

Appendix A in fact contains an evaluation like what is requested by reviewer #2.  Consider for example the first row.  It shows that translating each link grammar "A" edge to a leftward CoNLL "NMOD" edge would indeed be successful: this matches the CoNLL direction and label in 99% of the cases when CoNLL has an edge there at all (which is true for 86% of the 426 "A" links).

Scanning the last 3 columns of this table gives a clear picture of where the directed link parses line up with the CoNLL parses.  We could have artificially summarized the Appendix A table using one or two numbers.  However, we felt that the real information was in the line-by-line (type-by-type) pattern, in particular the fact that most link grammar label types mapped cleanly to a single direction, and that in most cases this agreed with the CoNLL direction and corresponded strongly to a single CoNLL label.  We also provided 100 actual parses in Appendix B so that researchers could see further whether our DAG parses would suit their research needs.

Note, however, that we are not holding out the converted link grammar parses as an ideal syntactic representation.  Indeed, we strongly believe that the "real" way to work on DAG parsers would be to produce an annotation scheme and a gold annotated resource.  Lacking one, the questions that we aimed to answer here were:

* Our experimental hypothesis was that the link grammar writers mentally associated a consistent direction with each link type, even though they did not annotate it.  We found that we could indeed recover consistent directions that satisfy the DAG constraints.  The first column of Appendix A can be directly applied to transform link parses into DAG parses without re-running the ILP.

* Are the converted link grammar parses good enough that people could use them to test PARSING ALGORITHMS?  This does not require that the parses agree with other annotation schemes (or even that they are linguistically defensible).  To experiment with the speed and accuracy of algorithms, one only needs a PLAUSIBLE distribution of data for train/test, which mimics the kind of data that one might have in a gold standard.  This is what we provide.  The point of the supplementary material is to demonstrate the plausibility.

* There has been a lot of ongoing development on link grammars for English and other languages, but it has been almost entirely outside the ACL community.  Is there a way to bring this work into the fold?  As far as we know, this is the first attempt.  These parses may serve as a source of extra features for other prediction or learning tasks (we actually had particular uses in mind that motivated the work).

* We pioneer the idea of using ILP to enhance incomplete annotations, using the ILP constraints to enforce internal consistency as well as formal requirements or linguistic preferences.

To Reviewer #3: There are no multi-planar link grammars at present.  If there were, yes, our ILP method would immediately extend.  In the meantime, one could do research on multi-planar DAG parsing by overlaying our DAG parses on the CoNLL trees (2 planes).

Thanks very much to Reviewer #1 for the additional citations and other suggestions.  (Regarding "converting corpora from other formats such as HPSG," we were referring to the work cited in the previous sentence.)

### 2014-07-10 doc/reviews-response.txt

Our SUPPLEMENTARY MATERIAL speaks to your evaluation concerns!

Reviewer #2: The last 3 columns of Appendix A show how the directed link parses line up with the CoNLL parses.  Most link grammar label types map cleanly to a single direction.  Usually this agrees with the CoNLL direction and corresponds strongly to a single CoNLL label.  We also provided 100 actual parses in Appendix B so that researchers could see further whether our DAG parses would suit their research needs.

(Row 1 shows that translating each link grammar "A" edge to a leftward CoNLL "NMOD" edge would match the CoNLL direction+label in 99% of the cases when CoNLL has a corresponding edge; that's 365 of 426 cases.)

But agreement with CoNLL wasn't our goal!  Our questions:

* Can we help people working on parsing algorithms?  Apparently yes.  We have produced a multi-headed projective DAG corpus.  The parses aren't gold, might not agree with other annotation schemes, or be linguistically defensible.  But they still give a GENERALLY PLAUSIBLE TRAINING/TEST SET for experimenting with the speed/accuracy of DAG parsing ALGORITHMS.  Someone should then build a gold corpus.

* Can ILP be used to enhance incomplete annotations?  Yes: we show how to enforce internal consistency along with well-formedness requirements.

* Did the link grammar writers have a link direction "in mind" even though it's not annotated?  Apparently yes.  In Appendix A we supply the solution.

* Can we ACLers make any use of the ignored labor of the link grammar writers?  Maybe.  These directed parses may serve as a source of extra features for other tasks.

Reviewer #3: No multi-planar link grammars exist at present.  If there were, our ILP method would immediately extend.  In the meantime, one can do research on 2-planar DAG parsing by overlaying our DAG parses on the CoNLL trees.

Reviewer #1: Thanks for the citations and suggestions.  We like the suggestions of other semantic dependency representations to compare (perhaps non-planar, disconnected, or not-manually-checked).

### 2014-09-21 doc/LinksToDAG.tex
Updated figures after rerunning the entire project.
commented out the call to the supplementary example parses in LinksToDAG.tex.
moved section 7 into figures with captions augmented conclusions some other minor edits

\section{Experiments and Results}
\begin{figure*}[ht!]
  \vspace{-6pt}
  \centering
  \input{figure/sentences.tikz}
  \caption{\small Example Parses: The blue upper edges show CoNLL dependency parses; the red lower edges show oriented link grammar parses. Vertical edges are those with parent 0.  See Appendix~\ref{app:parses} for formatting details and more examples.}
  \label{fig:parses}
  \vspace{-6pt}
\end{figure*}

We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.  Our largest run took 1.5 hours. \noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  Detailed results from this run are included as Appendix~\ref{app:linktypes}.  On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multiheaded word.
%\noteJE{2 significant digits is enough}
%\noteJE{update these numbers}

\subsection{Stability of Results}\label{sec:stability}
\begin{figure}[ht!]
  \small
  \centering
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{Convergence to the direction mapping obtained on the largest run.  Precision relative to that run is very high even for small datasets.  Recall grows as more link types are seen.}\label{fig:convergence}
\vspace{-6pt}
\end{figure}

We worried that the direction mapping might be unstable and sensitive to the input corpus.  Figure~\ref{fig:convergence} shows otherwise (for both English and Russian). \noteJE{IMPORTANT! current graph is with fine-grained tags; rerun with the coarse-grained tags as in the rest of the paper} Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.  The only disadvantage to using small data is low recall relative to the large run---many of the labels $L$ are not observed yet and so we do not yet allow either direction ($r_L=\ell_L=0$).

@@

The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in Appendix~\ref{app:linktypes}, and the actual parses are contrasted in Figure~\ref{fig:parses} and Appendix~\ref{app:parses}.

@@

More generally, ILP may be a valuable technology for enriching existing annotated corpora.  For example, the Penn Treebank \cite{TREEBANK-1993} omitted types of annotations that plausibly could be reconstructed by rule.

% \newpage
\bibliographystyle{acl2014}
\bibliography{LinksToDAG}

>>>>

\section{Experiments and Results}

\begin{figure*}[ht!]
  \vspace{-6pt}
  \centering
  \input{figure/sentences.tikz}
  \input{figure/sample_parses.tikz}
  \caption{\small The blue upper edges show CoNLL dependency parses; the red lower edges show our oriented version of the link grammar parses. Edges are shown as dotted lines if they appear only in one parse; these are highlighted in orange if the child has multiple parents.  Edges that appear in both parses are solid lines, drawn thicker if the directions do not match.  Vertical edges have parent 0.  For 100 example parses, see Appendix~\ref{app:parses} at \url{http://tidyurl.com/LinksToDAG_supplement}.}
  \label{fig:parses}
  \vspace{-6pt}
\end{figure*}

\begin{table*}[ht]
% \renewcommand{\arraystretch}{1.2}    % more space between rows
\input{figure/sample_link_analysis_coarse_table.tex}
\caption{Our solution, i.e., our reconstruction of the ``intended'' direction for each link type in the English Link Grammar.  We also indicate the extent to which each of these link types (1) has a single dominant direction, (2) participates in multi-headed constructions, and (3) corresponds to CoNLL links of a predictable direction and type.  For space reasons, we show only the start of this table---the full table can be found in Appendix~\ref{app:linktypes} at \url{http://tidyurl.com/LinksToDAG_supplement}.}\label{tab:linktypes}
\end{table*}

\begin{figure}[t!]
  \small
  \centering
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{Convergence to the direction mapping obtained on the largest run.  Precision relative to that run is very high even for small datasets.  Recall grows as more link types are seen.}\label{fig:convergence}
\vspace{-6pt}
\end{figure}

We solved our ILP problem using the SCIP Optimization Suite \cite{achterberg2009scip}, encoding it using the ZIMPL language \cite{Koch2004}.  Our largest run took 1.5 hours. \noteJH{A run of 10,000 sentences takes 3142 seconds (update in final version!)}  On English, only \input{figure/allowed_labels_both.tex}of \input{figure/allowed_labels_total.tex}link {\em types} allowed both directions, and only $\sum_{ij} s_{ij} = \input{figure/label_token_disallowed}$ of \input{figure/label_token_total.tex}link {\em tokens} required a disallowed direction via slack.
\input{figure/conll_analysis_sentences_multiheaded.tex}of the English sentences but (alas) only \input{figure/russian_multiheaded.tex}of the Russian ones had at least one multiheaded word.  See Table~\ref{tab:linktypes} and 
Appendix~\ref{app:linktypes} for detailed results.
%\noteJE{2 significant digits is enough}
%\noteJE{update these numbers}

\subsection{Stability of Results}\label{sec:stability}

We worried that the direction mapping might be unstable and sensitive to the input corpus.  Happily, Figure~\ref{fig:convergence} shows otherwise (for both English and Russian). \noteJE{IMPORTANT! current graph is with fine-grained tags; rerun with the coarse-grained tags as in the rest of the paper} Using even a small prefix of the data got very high-precision results, in the sense that nearly all $r_L$ or $\ell_L$ variables that were 1 in this lightly trained mapping were also 1 in our largest run.  The only disadvantage to using small data is low recall relative to the large run---many of the labels $L$ are not observed yet and so we do not yet allow either direction ($r_L=\ell_L=0$).

@@

The resulting English corpus uses a syntactic annotation scheme that is somewhat different from the CoNLL annotations.  Differences are tabulated in
Appendix~\ref{app:linktypes}, while the actual parses are contrasted in Appendix~\ref{app:parses}.  By permission of the program chairs, these 
appendices are available at the anonymous URL \url{http://tidyurl.com/LinksToDAG_supplement}.  Fragments are shown in 
Table~\ref{tab:linktypes} and Figure~\ref{fig:parses}.

@@

ILP may {\em in general} be a valuable technology for enriching existing annotated corpora.  For example, the Penn Treebank project \cite{TREEBANK-1993} deliberately omitted types of annotations that plausibly could be added automatically.  ILP can help by leveraging unsupervised corpus-wide information \cite{ravi2009}, enforcing annotations that are simultaneously well-formed per sentence and consistent across sentences.

% \subsection{Sample of English Parses}
% 
% \begin{figure*}[ht!]
%   \vspace{-6pt}
%   \centering
%   \input{figure/sample_parses.tikz}
%   \vspace{-6pt}
% \end{figure*}

%\twocolumn

\newpage
\bibliographystyle{acl2014}
\bibliography{LinksToDAG}
